Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2019-01-17 scagnostics.Rmd
done with red chunks

```{r bbhGuided, results="hide", fig.cap="Projections corresponding to the maxima of three indices: TIC, splines2D and 1-convex. Projections a, b found by TIC and splines2d are very similar, and involve the same three parameters, ra, dec and time. The 1-convex index finds a very different view.", dev = "png", dpi=300}
bbhD <- read_csv("data/posterior_samples.csv")
bbhM <- rescale(select(bbhD,-m2, -chi_eff))
if(!file.exists("cache/bbhGuidedTIC.rda")){
  set.seed(2018)
  bbhTour1 <- save_history(bbhM, guided_tour(mineIndex("TIC")))
  bbhTourFull1 <- as.list(interpolate(bbhTour1))
  save(bbhTour1, bbhTourFull1, file = "cache/bbhGuidedTIC.rda")
} else {
  load("cache/bbhGuidedTIC.rda")
}
if(!file.exists("cache/bbhGuidedSpline.rda")){
  set.seed(2018)
  bbhTour2 <- save_history(bbhM, guided_tour(splineIndex()))
  bbhTourFull2 <- as.list(interpolate(bbhTour2))
  save(bbhTour2, bbhTourFull2, file = "cache/bbhGuidedSpline.rda")
} else {
  load("cache/bbhGuidedSpline.rda")
}
if(!file.exists("cache/bbhGuidedConvex.rda")){
  set.seed(2018)
  bbhTour3 <- save_history(bbhM, guided_tour(invConvexIndex()))
  bbhTourFull3 <- as.list(interpolate(bbhTour3))
  save(bbhTour3, bbhTourFull3, file = "cache/bbhGuidedConvex.rda")
} else {
  load("cache/bbhGuidedConvex.rda")
}

@@

```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", dev = "png", dpi=300}
bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
if(!file.exists("cache/bbhGuidedTIC2.rda")){
  set.seed(2018)
  bbhTour12 <- save_history(bbhM2, guided_tour(mineIndex("TIC")))
  bbhTourFull12 <- as.list(interpolate(bbhTour12))
  save(bbhTour12, bbhTourFull12, file = "cache/bbhGuidedTIC2.rda")
} else {
  load("cache/bbhGuidedTIC2.rda")
}
if(!file.exists("cache/bbhGuidedSpline2.rda")){
  set.seed(2018)
  bbhTour22 <- save_history(bbhM2, guided_tour(splineIndex()))
  bbhTourFull22 <- as.list(interpolate(bbhTour22))
  save(bbhTour22, bbhTourFull22, file = "cache/bbhGuidedSpline2.rda")
} else {
  load("cache/bbhGuidedSpline2.rda")
}
if(!file.exists("cache/bbhGuidedConvex2.rda")){
  set.seed(2018)
  bbhTour32 <- save_history(bbhM2, guided_tour(invConvexIndex()))
  bbhTourFull32 <- as.list(interpolate(bbhTour32))
  save(bbhTour32, bbhTourFull32, file = "cache/bbhGuidedConvex2.rda")
} else {
  load("cache/bbhGuidedConvex2.rda")
}

@@

Figure \ref{fig:bbhGuided3} shows six maxima obtained by different starts, for two types of parameters: first, spin related parameters (i.e. alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e. ra, dec and distance). Four of the six (a-d) are almost identical, but not interesting projections. Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.  \textcolor{red}{While none of these projections are particularly interesting on their own, moving between them can be revealing.}

Choosing a different subset of variables reveals something new. The subspace of m1, ra, chi\_tot, alpha, distance, dec produces a more refined view of Figure \ref{fig:bbhGuided3} projection f. When alpha contributes in contrast to dec, the relationship between the points is almost perfectly on a curve. This is shown in Figure \ref{fig:constructedExample}. Manually reconstructing this can be done by differencing the two parameters, in their original units. \textcolor{red}{This result is highlighting the importance of improved optimisation, that would allow to identify this relation given the full list of input parameters.}

@@

```{r constructedExample, results="hide", fig.cap="Selecting subset of 6 parameters, for which spline index can be used to improve constraint on parameter combination wrt ra vs dec.", dev = "png", dpi=300}
bbhC <- rescale(select(bbhD, m1, ra, chi_tot, alpha, distance, dec))
if(!file.exists("cache/constructedExample.rda")){
  set.seed(1999)
  tC <- save_history(bbhC, guided_tour(splineIndex()))
  save(tC, file = "cache/constructedExample.rda")
} else {
  load("cache/constructedExample.rda")
}
iL <- length(as.list(tC))
dProj <- bbhC %*% as.list(tC)[[iL]]
fProj <- as.list(tC)[[iL]]
colnames(fProj) <- c("PP1", "PP2")
colnames(dProj) <- c("PP1", "PP2")
fProj <- as_tibble(fProj) %>%
  add_column(label=colnames(bbhC))
p1 <- ggplot(as_tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
  geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y), color="grey") + 
  xlim(-0.9, 0.1) + ylim(-0.1, 1.1) +
  geom_segment(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
  geom_text(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
p2 <- ggplot(as_tibble(bbhD),aes(dec-alpha, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
p3 <- ggplot(as_tibble(bbhD),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
grid.arrange(p1, p2, p3, ncol=3)
```

>>>>

```{r bbhGuided, results="hide", fig.cap="Projections corresponding to the maxima of three indices: TIC, splines2D and 1-convex. Projections a, b found by TIC and splines2d are very similar, and involve the same three parameters, ra, dec and time. The 1-convex index finds a very different view.", fig.height = 3.2, dev = "png", dpi=300}
bbhD <- read_csv("data/posterior_samples.csv")
bbhM <- rescale(select(bbhD,-m2, -chi_eff))
if(!file.exists("cache/bbhGuidedTIC.rda")){
  set.seed(2018)
  bbhTour1 <- save_history(bbhM, guided_tour(mineIndex("TIC")))
  bbhTourFull1 <- as.list(interpolate(bbhTour1))
  save(bbhTour1, bbhTourFull1, file = "cache/bbhGuidedTIC.rda")
} else {
  load("cache/bbhGuidedTIC.rda")
}
if(!file.exists("cache/bbhGuidedSpline.rda")){
  set.seed(2018)
  bbhTour2 <- save_history(bbhM, guided_tour(splineIndex()))
  bbhTourFull2 <- as.list(interpolate(bbhTour2))
  save(bbhTour2, bbhTourFull2, file = "cache/bbhGuidedSpline.rda")
} else {
  load("cache/bbhGuidedSpline.rda")
}
if(!file.exists("cache/bbhGuidedConvex.rda")){
  set.seed(2018)
  bbhTour3 <- save_history(bbhM, guided_tour(invConvexIndex()))
  bbhTourFull3 <- as.list(interpolate(bbhTour3))
  save(bbhTour3, bbhTourFull3, file = "cache/bbhGuidedConvex.rda")
} else {
  load("cache/bbhGuidedConvex.rda")
}

@@

```{r bbhGuided2, results="hide", fig.cap="Projections of the reduced 8D space corresponding to the maxima of three indices: TIC, splines2d and 1-convex.", fig.height = 3.2, dev = "png", dpi=300}
bbhM2 <- rescale(select(bbhD,-m2, -chi_eff, -time, -dec, -ra))
if(!file.exists("cache/bbhGuidedTIC2.rda")){
  set.seed(2018)
  bbhTour12 <- save_history(bbhM2, guided_tour(mineIndex("TIC")))
  bbhTourFull12 <- as.list(interpolate(bbhTour12))
  save(bbhTour12, bbhTourFull12, file = "cache/bbhGuidedTIC2.rda")
} else {
  load("cache/bbhGuidedTIC2.rda")
}
if(!file.exists("cache/bbhGuidedSpline2.rda")){
  set.seed(2018)
  bbhTour22 <- save_history(bbhM2, guided_tour(splineIndex()))
  bbhTourFull22 <- as.list(interpolate(bbhTour22))
  save(bbhTour22, bbhTourFull22, file = "cache/bbhGuidedSpline2.rda")
} else {
  load("cache/bbhGuidedSpline2.rda")
}
if(!file.exists("cache/bbhGuidedConvex2.rda")){
  set.seed(2018)
  bbhTour32 <- save_history(bbhM2, guided_tour(invConvexIndex()))
  bbhTourFull32 <- as.list(interpolate(bbhTour32))
  save(bbhTour32, bbhTourFull32, file = "cache/bbhGuidedConvex2.rda")
} else {
  load("cache/bbhGuidedConvex2.rda")
}

@@

Figure \ref{fig:bbhGuided3} shows six maxima obtained by different starts, for two types of parameters: first, spin related parameters (i.e. alpha, theta\_jn, chi\_tot and chi\_p), and second position related parameters (i.e. ra, dec and distance). Four of the six (a-d) are almost identical, but not interesting projections. Projection f has the highest PP index value but it is primarily the view seen in the bivariate plot of dec and ra.  While none of these projections are particularly interesting on their own, moving between them can be revealing.

Choosing a different subset of variables reveals something new. The subspace of m1, ra, chi\_tot, alpha, distance, dec produces a more refined view of Figure \ref{fig:bbhGuided3} projection f. When alpha contributes in contrast to dec, the relationship between the points is almost perfectly on a curve. This is shown in Figure \ref{fig:constructedExample}. Manually reconstructing the optimal projection (left plot) can be done by differencing the two parameters, in their original units. This highlights the importance of improved optimisation, that would use tiny final steps to polish the view to a finer optimal projection and possibly remove noise induced by small contributions of many variables.

@@

```{r constructedExample, results="hide", fig.cap="Manual reconstruction of an optimal projection (left), constructed by differencing alpha from dec in the original units against ra (middle), compared with the two main variables (right).", fig.height = 3.2, dev = "png", dpi=300}
bbhC <- rescale(select(bbhD, m1, ra, chi_tot, alpha, distance, dec))
if(!file.exists("cache/constructedExample.rda")){
  set.seed(1999)
  tC <- save_history(bbhC, guided_tour(splineIndex()))
  save(tC, file = "cache/constructedExample.rda")
} else {
  load("cache/constructedExample.rda")
}
iL <- length(as.list(tC))
dProj <- bbhC %*% as.list(tC)[[iL]]
fProj <- as.list(tC)[[iL]]
colnames(fProj) <- c("PP1", "PP2")
colnames(dProj) <- c("PP1", "PP2")
fProj <- as_tibble(fProj) %>%
  add_column(label=colnames(bbhC))
p1 <- ggplot(as_tibble(dProj), aes(PP1, PP2))+geom_point(alpha=0.02) + theme(aspect.ratio=1)+ 
  geom_path(data=getCircle(-0.7, 0.6, 0.2, 0.24), aes(x=x, y=y), color="grey") + 
  xlim(-0.9, 0.1) + ylim(-0.1, 1.1) +
  geom_segment(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x1, xend=x2, y=y1, yend=y2)) +
  geom_text(data=getAxes(fProj, -0.7, 0.6, 0.2, 0.24), aes(x=x2, y=y2, label=label), size=2.5)
p2 <- ggplot(as_tibble(bbhD),aes(dec-alpha, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
p3 <- ggplot(as_tibble(bbhD),aes(dec, ra)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)
grid.arrange(p1, p2, p3, ncol=3)
```
